---
title: "SHAirData"
output: html_document
date: "2024-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(ggplot2)
library(rio)
library(ggExtra)
```

## Analyte omissions - air and wristbands

# benz[a]anthracene
# benz[j] and [e]aceanthrylene
# 2,6-dimethylnaphthalene
# 2,6-diethylnaphthalene

## Potential sample issues 
# SH004_A_04
# SH003_WB_08
# SH007_A_08

## Reccommend Sensitivity Analysis 
### SH019_A_03
### SH007_A_06

## Omit
### SH044_WB_08
### SH045_WB_08


## Load in all data from the deidentified data book copy in the share drive

```{r}
Data_list <- import_list("SHAirData_Deidentified.xlsx")
```

## Rename each sheet as individual dataframes

```{r}
monthly_meta <- Data_list$monthly_meta
part <- Data_list$participants
comm_coord <- Data_list$community_samples
WBD <- Data_list$chemical_data_WB
AirD <- Data_list$chemical_data_Air
EnvD <- Data_list$environmental_data
chem_info <- Data_list$chemical_parameter
```

## Cleaning up and annotating the air data

```{r}
Air_clean <- AirD %>%
  mutate(Deployed_by = ifelse(str_detect(Name, c("SH0")), "Participant", ifelse(str_detect(Name, "SMS_I|SHS"), "Teacher", "OSU"))) %>%
  separate(Name, c("ID","X","Y")) %>% #splitting name to pull out ID and month; some samples have extra annotations for sample type (A) or outdoor(O)/indoor(I)
  mutate(Indoor = ifelse(X == "I", "yes", "no")) %>% #identify indoor classroom sample (to omit from outdoor air analyses)
  mutate(Month = ifelse(is.na(Y), X, Y)) %>% #extracting the month
  mutate(Type = "A") %>% #all of the samples in this dataset are Air samples (A), opposed to wristbands (WB)
  filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  select(ID, Type, Month, ParameterName, Result, Flag, Indoor, Deployed_by) %>%
  mutate(Flag = replace_na(Flag, "D")) #D represents detections without J or B flags; U-flags were below detection limits

# Samples collected per time point:
Air_clean %>%
  select(ID, Month, Deployed_by) %>%
  distinct() %>%
  group_by(Month, Deployed_by) %>%
  summarize(n()) %>%
  pivot_wider(names_from = Deployed_by, values_from = 'n()')


##Visualization of Air Data (Alison)


```{r}
# concentration distributions by chemical and month
Air_clean %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(x = "Concentration ng/m^3", y = "Chemical Detected")
```
```{r}
# concentration distributions by chemical and month
Air_clean2 <- Air_clean %>% filter(ID != "NW10",ID != "SW10", ID != "C" )

Air_clean2 %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle=90)) +
  labs(x = "Concentration ng/m^3", y = "Chemical Detected") 
```

```{r}
# Units are ng/m^3
# median heatmap
Air_clean1 <- Air_clean %>% filter(Month != "01")
Air_clean1 %>%
  group_by(Month, ParameterName) %>%
  summarize(Median = median(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Median)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed() +
  labs(x = "Month", y = "Chemical Parameter Median Value ng/m^3")+
  theme(axis.text.x = element_text(angle=90))

# mean heatmap
Air_clean1 %>%
  group_by(Month, ParameterName) %>%
  summarize(Mean = mean(Result), Median = median(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Mean)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()+
  labs(x = "Month", y = "Chemical Parameter Mean Value ng/m^3")+
  theme(axis.text.x = element_text(angle=90))

# detection frequency
Air_clean1 %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), Median = median(Result), n = n()) %>%
  mutate(Detection_Frequency = n/max(n)) %>% #calculate detection frequency
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Detection_Frequency)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c() +
  coord_fixed()+
  labs(x = "Month", y = "Chemical Parameter Detection Frequency")+
  theme(axis.text.x = element_text(angle=90))

#Omitting Creosote Site Samplers

Air_clean_nosite <- Air_clean2 %>% filter(Month != "01")
Air_clean_nosite %>%
  group_by(Month, ParameterName) %>%
  summarize(Median = median(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Median)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed() +
  labs(x = "Month", y = "Chemical Parameter Median Value ng/m^3")+
  theme(axis.text.x = element_text(angle=90))

# mean heatmap
Air_clean_nosite %>%
  group_by(Month, ParameterName) %>%
  summarize(Mean = mean(Result), Median = median(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Mean)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()+
  labs(x = "Month", y = "Chemical Parameter Mean Value ng/m^3")+
  theme(axis.text.x = element_text(angle=90))

# detection frequency
Air_clean_nosite %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), Median = median(Result), n = n()) %>%
  mutate(Detection_Frequency = n/max(n)) %>% #calculate detection frequency
  ggplot(aes(y=reorder(ParameterName, Median), x=Month, fill = Detection_Frequency)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c() +
  coord_fixed()+
  labs(x = "Month", y = "Chemical Parameter Detection Frequency")+
  theme(axis.text.x = element_text(angle=90))
```
