---
title: "St. Helens Data"
output: html_document
date: "2024-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(ggplot2)
library(rio)
library(ggExtra)
library(ggcorrplot)
library(corrplot)
library(psych)
library(ggbreak)
library(patchwork)
library(EnvStats)
library(magrittr)
library(Tplyr)
library(knitr)
```

# Log of changes 

# 1/27 to 2/12/2024:
## changed order of the heatmaps to be consistent (all by det freq) 
## corrected the list of analytes used in the correlation plot to ensure there are 5 or more pairs of observations per square
## created a variety of plots using coefficients of variation with at least 10 and 30 detections

# 2/12 - 2/26/2024
## added coefficient of variation colors to indicate number of detections considered

# 2/26 - 3/26/2024
## updated figures included in dissertation


# Analyte omissions

## benz[a]anthracene
## benz[j] and [e]aceanthrylene
## 2,6-dimethylnaphthalene
## 2,6-diethylnaphthalene

# Problematic samples

## Reccommend Sensitivity Analysis
### SH019_A_03
### SH007_A_06

## Omit
### SH044_WB_08
### SH045_WB_08


# Load in all data from the deidentified data book 

```{r}
Data_list <- import_list("SHAirData_Deidentified.xlsx")
```

# Rename each sheet as individual dataframes

```{r}
monthly_meta <- Data_list$monthly_meta
part <- Data_list$participants
comm_coord <- Data_list$community_samples
WBD <- Data_list$chemical_data_WB
AirD <- Data_list$chemical_data_Air
EnvD <- Data_list$environmental_data
```

# Clean Up WB Data

```{r}
#WB data currently in ng/g
(WB_clean <- WBD %>%
  filter(Name != "SH044_WB_08",
         Name != "SH045_WB_08") %>% #samples to omit
   filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  separate(Name, into = c("ID", "Type", "Month")) %>%
  select(ID, Month, ParameterName, Result, Flag) %>%
  mutate(Flag = replace_na(Flag, "D")) %>% #D represents detections; easier to work with than NA values
  filter(Flag != "U"))
```

# Data Visualization - data distributions and heatmaps

```{r}
# concentration distributions by chemical and month
WB_clean %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() 

# median ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(median = median(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = median)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed() 

# mean ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = mean)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()

# detection frequency
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = det_freq)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c() +
  coord_fixed()

```

# Stacked bar chart - median concentrations (>50% rate of detection within each month)

```{r}
#stacked bar chart; limit to 50% detection rate
#23 chemicals detected in at least 50%

pal <- c("#5A5156", "#E4E1E3", "#1C7F93", "#1CBE4F", "#B5EFB5", 
           "#3283FE", "#FEAF16", "#B00068", "#1CFFCE", "#90AD1C", 
           "#DEA0FD", "#F8A19F", "#325A9B", "#C4451C", "#1C8356", 
           "#85660D", "#FBE426", "#683B79", "#822E1C", 
           "#F7E1A0", "#C075A6", "#BDCDFF", 
           "#7ED7D1", "#2ED9FF", "#66B0FF", "#3B00FB")

# stacked means (ng/g)
WB_clean %>% #this is mean concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>% #at least 50% detection rate
  ggplot(aes(x=Month, y= mean, fill = reorder(ParameterName, mean))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 

# stacked medians (ng/g)
WB_clean %>% #this is median concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), median = median(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>%
  ggplot(aes(x=Month, y= median, fill = reorder(ParameterName, median))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 
```

# Proportion plot based on median concentrations

```{r}
#chemicals in 50% across all months - median
p1 <- WB_fifty_percent_dets %>%  
  group_by(Month, ParameterName) %>%
  summarize(median = median(Result), n = n(), mean = mean(log10(Result))) %>%
  mutate(det_freq = n/max(n), prop = median/sum(median)) %>%
  filter(det_freq > .5) %>%
  ggplot(aes(x= as.factor(Month), y= prop, color = fct_rev(reorder(ParameterName, prop)), group = ParameterName)) +
  geom_point(aes(size = n)) +
  geom_line() +
  scale_color_discrete(type  = pal_dk) +
    scale_size_continuous(range=c(0.5,5)) +
  theme_bw() +
    theme(text=element_text(size=20))

p2 <- p1 + scale_y_break(c(0.11, 0.15), scales = 0.6, ticklabels=c(0.15, 0.18, 0.21)) + 
  xlab(NULL) + ylab(NULL) 

p2
```

# Correlation of chemicals

```{r}
# First convert data to a matrix with "NA" values for BLOD instances where each row is an analyte (5+ detections) and each row is a different WB sample (at least one detection).

(five_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>4)) 

chemical_list <- as.vector(five_dets$ParameterName) # save the list of analytes with 5 or more detections (45 analytes)
WB_clean_five_dets <- subset(WB_clean, ParameterName %in% chemical_list) # create a new WB dataset with only the list of analytes saved above

WB_clean_wide <- WB_clean_five_dets %>%
  select(-Flag) %>%
  pivot_wider(names_from = ParameterName, values_from = Result)

WB_clean_wide <- as.data.frame(WB_clean_wide)

row.names(WB_clean_wide) <- paste(WB_clean_wide$ID, WB_clean_wide$Month, sep="_")

WB_mat_ready <- WB_clean_wide %>%
  select(-ID, -Month)

(WB_mat <- as.matrix(WB_mat_ready))

# Identify "problem" chemicals that have less than 5 pairs of data for a comparison

problems <- pairwiseReport(WB_mat, diagonal = F, cut = 5)
print(problems, short=F)
#Problems output shows those chemicals that have less than 5 pairs for a comparison

problem_list_rows <- c("2-methylanthracene", "6-methylchrysene", "anthracene", "benzo[a]fluorene", "benzo[a]pyrene", "benzo[b]fluorene", "benzo[ghi]perylene", "benzo[j]fluoranthene", "benzo[k]fluoranthene", "coronene", "cyclopenta[cd]pyrene", "indeno[1,2,3-cd]pyrene", "perylene")

#removing chemicals with fewer than 5 pairs of observations from the matrix
(five_pairs <- WB_clean_five_dets %>%
    filter(!ParameterName %in% problem_list_rows)) 

# log10 transform data and shape into a matrix for correlation plot
(logWB_clean_wide <- five_pairs %>%
  mutate(log10Result = log10(Result)) %>% 
  select(-Flag, -Result) %>%
  pivot_wider(names_from = ParameterName, values_from = log10Result))

logWB_clean_wide <- as.data.frame(logWB_clean_wide)
row.names(logWB_clean_wide) <- paste(logWB_clean_wide$ID, logWB_clean_wide$Month, sep="_")
(logWB_mat_ready <- logWB_clean_wide %>%
  select(-ID, -Month))

(logWB_mat <- as.matrix(logWB_mat_ready))

# Create a correlation matrix and plot

logWB_chem_corr <- cor(logWB_mat, use="pairwise.complete.obs")
logWB_pmat <- cor_pmat(logWB_mat,
  method = "spearman",
  alternative = "two.sided",
  conf.level = 0.95)

ggcorrplot(logWB_chem_corr, type = "lower") #simple

#Version used in dissertation; categories of PAHs added in PPT (retene, 2-ring PAHs, 3+ ring PAHs, 3-ring PAHs)
corrplot(logWB_chem_corr, order = 'hclust', addrect = 4, tl.col = 'black', cl.ratio = 0.2, tl.srt = 45, tl.cex = 0.9, p.mat = logWB_pmat, insig = 'blank', tl.offset = 2)

```

# Timeseries number of detections

```{r}
WB_cleangrp <- WB_clean %>%
  mutate(Group = ifelse(Month %in% c(2,4,6,8), "1", "2"))

WB_cleangrp %>% # group 1
  filter(Group == 1) %>%
  group_by(ID, Month) %>%
  summarize(sumPAH = sum(Result), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = nPAH, color = ID)) +
  geom_point() +
  geom_path() +
  theme_bw()

WB_cleangrp %>% # group 2
  filter(Group == 2) %>%
  group_by(ID, Month) %>%
  summarize(sumPAH = sum(Result), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = nPAH, color = ID)) +
  geom_point() +
  geom_path() +
  theme_bw()
```

## Timeseries concentrations - played with grouping by individual and different chemical categories

```{r}
# group by individual; G1
WB_cheminfo %>% 
  filter(Group == 1) %>%
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings,"A"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = Type)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(ID), ncol = 4, scales = "free") +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures \n for Group 1 Faceted by ID")  +
  theme_bw()

# group by individual; G2
WB_cheminfo %>% 
  filter(Group == 2) %>%
  filter(ID != "SH002", 
         ID != "SH016",
         ID != "SH032") %>%  #remove SH002, SH016, and SH032 (only one observation)
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings,"A"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = Type)) +
  geom_point() +
  geom_path() +
  facet_wrap(vars(ID), ncol = 4, scales = "free") +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures \n for Group 2 Faceted by ID") +
  theme_bw()

# create timeseries plots for groups of chemicals

(WB_Name <- WB_cleangrp %>% #create names to match the chem_info data
  rename("Name" = ParameterName))

(WB_cheminfo <- left_join(WB_Name, chem_info, by ="Name")) 

WB_cheminfo %>% # group 1 - facet by rings (log10 transformed)
  filter(Group == 1) %>%
  group_by(ID, Month, Rings) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Rings) +
  theme_bw()

WB_cheminfo %>% # group 2 - facet by rings (log10 transformed)
  filter(Group == 2) %>%
  group_by(ID, Month, Rings) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Rings) +
  theme_bw()

WB_cheminfo %>% # all - facet by rings (log10 transformed) 
  group_by(ID, Month, Rings) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Rings) +
  theme_bw()

WB_cheminfo %>% # group 1 - facet by rings (log10 transformed); combine ring sized 4-7
  filter(Group == 1) %>%
  mutate(Size = ifelse(Rings == 2, Rings, ifelse(Rings == 3, Rings, "4+"))) %>%
  group_by(ID, Month, Size) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Size) +
  labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures Faceted by Size (# of Rings)") +
  theme_bw()

WB_cheminfo %>% # group 2 - facet by rings (log10 transformed); combine ring sized 4-7
  filter(Group == 2) %>%
  mutate(Size = ifelse(Rings == 2, Rings, ifelse(Rings == 3, Rings, "4+"))) %>%
  group_by(ID, Month, Size) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Size) +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures Faceted by Size (# of Rings)") +
  theme_bw()

WB_cheminfo %>% # all - facet by rings (log10 transformed); combine ring sized 4-7
  mutate(Size = ifelse(Rings == 2, Rings, ifelse(Rings == 3, Rings, "4+"))) %>%
  group_by(ID, Month, Size) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point() +
  geom_path() +
  facet_wrap(~ Size) +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures Faceted by Size (# of Rings)") +
  theme_bw()

WB_cheminfo %>% # facet by group and rings (log10 transformed); combine ring sized 4-7; hide ID
  mutate(Size = ifelse(Rings == 2, Rings, ifelse(Rings == 3, Rings, "4+"))) %>%
  group_by(ID, Month, Size, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point(show.legend = FALSE) +
  geom_path(show.legend = FALSE) +
  facet_grid(rows = vars(Size), cols = vars(Group)) +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures Faceted by Sampling Group (Cols) and # of Rings (Rows)") +
  theme_bw()

WB_cheminfo %>% # facet by group and rings + alkylation (log10 transformed)
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings,"A"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result))) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point(show.legend = FALSE) +
  geom_path(show.legend = FALSE) +
  facet_grid(rows = vars(Type), cols = vars(Group), scales = "free") +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures \n Faceted by Sampling Group (Cols) and # of Rings + Alkylation (Rows)") +
  theme_bw()

WB_cheminfo %>% # facet by group and rings+ alkylation (log10 transformed); no SH018
  filter(ID != "SH018") %>%
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings,"A"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = Month, y = sumPAH, color = ID)) +
  geom_point(show.legend = FALSE) +
  geom_path(show.legend = FALSE) +
  facet_grid(rows = vars(Type), cols = vars(Group), scales = "free") +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures \n Faceted by Sampling Group (Cols) and # of Rings + Alkylation (Rows)") +
  theme_bw()

WB_cheminfo %>% # all; just retene
  filter(Name == "retene") %>%
  ggplot(aes(x = Month, y = log10(Result), color = ID)) +
  geom_point(show.legend = FALSE) +
  geom_path(show.legend = FALSE) +
  facet_grid(cols = vars(Group), scales = "free") +
    labs(y = "Log10 Retene Concentration (ng/g)", 
       title = "Retene Concentrations of Repeated Measures Faceted by Sampling Group") +
  theme_bw()

```

## Timeseries boxplots - by number of rings and alkylation

```{r}
#WB_cheminfo created in chunk above

WB_cheminfo %>% # all - facet by rings (log10 transformed)
  group_by(ID, Month, Rings) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = as.factor(Month), y = sumPAH)) +
  geom_boxplot() +
  facet_wrap(~ Rings) +
  theme_bw()

WB_cheminfo %>% # all - facet by rings (log10 transformed); combine ring sized 4-7
  mutate(Size = ifelse(Rings == 2, Rings, ifelse(Rings == 3, Rings, "4+"))) %>%
  group_by(ID, Month, Size) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  ggplot(aes(x = as.factor(Month), y = sumPAH)) +
  geom_boxplot() +
  facet_wrap(~ Size) +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
       title = "Sum PAH Concentrations of Repeated Measures Faceted by Size (# of Rings)") +
  theme_bw()

WB_cheminfo$Month <-as.factor(WB_cheminfo$Month)
WB_cheminfo$Month <- recode(WB_cheminfo$Month, '2' = "Nov", '3' = "Dec", '4' = "Jan", '5' = "Feb", '6' = "Mar", '7' = "Apr", '8' = "May", '9' = "Jun")

p1 <- WB_cheminfo %>% # facet by group and rings+ alkylation (log10 transformed); COLOR ID
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings," Alkylated"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  filter(str_detect(Type, "2")) %>%
  ggplot(aes(x = Month, y = sumPAH)) +
  geom_boxplot() +
  facet_wrap(~Type, scales = "free", ncol = 2) +
    labs(x = "",
         y = element_blank()) +
  theme_bw() +
  theme(panel.spacing = unit(0.7, "cm")) 

p2 <- WB_cheminfo %>% # facet by group and rings+ alkylation (log10 transformed); COLOR ID
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings," Alkylated"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  filter(str_detect(Type, "3")) %>%
  ggplot(aes(x = Month, y = sumPAH)) +
  geom_boxplot() +
  facet_wrap(~Type, scales = "free", ncol = 2) +
    labs(y = "Log10 Sum PAH Concentration (ng/g)", 
         x = "") +
  theme_bw() +
  theme(panel.spacing = unit(0.7, "cm")) 

p4 <- WB_cheminfo %>% # facet by group and rings+ alkylation (log10 transformed); COLOR ID
  mutate(Type = ifelse(Alkylated == "yes", paste0(Rings," Alkylated"), ifelse(Rings >= 4, "4+", Rings))) %>%
  group_by(ID, Month, Type, Group) %>%
  summarize(sumPAH = log10(sum(Result)), nPAH = n()) %>% #sum concentration and number of detections
  filter(str_detect(Type, "4")) %>%
  ggplot(aes(x = Month, y = sumPAH)) +
  geom_boxplot() +
  facet_wrap(~factor(Type, levels=c('4+', '4 Alkylated')), scales = "free", ncol = 2) +
    labs(x = "Month of Sampling",
         y = element_blank()) +
  theme_bw() +
  theme(panel.spacing = unit(0.7, "cm")) 

WB_cheminfo$nicetitle = "3 Alkylated (Retene)"

p3 <- WB_cheminfo %>% # just retene
  filter(Name == "retene") %>%
  ggplot(aes(x =Month, y = log10(Result))) +
  geom_boxplot() +
  facet_wrap(~nicetitle) +
    labs(y = element_blank(),
         x = element_blank()) +
  theme_bw()

layout <- "
AA#
BBC
DD#
"
p1 + p2 + p3 + p4 + 
  plot_layout(design = layout)

```

# PCA Plot

```{r}
loadings <- read.csv("PCA_Loadings.csv")
scores <- read.csv("PCA_Scores.csv")

ggplot(scores, aes(PC1, PC2)) +
  geom_point() +
  theme_bw()
```

# Quick Look into high concentrations (looks like mostly in group 1)

```{r}
# 1. filter data by any concentrations above 500 and 1000 ng/g

WB_clean %>% 
  filter(Result > 1000) #SH018 in 02 and 04 - verify data, followed by SH048_06
WB_clean %>% 
  filter(Result > 500) #SH018 in 06 and 08; SH048_06 again; phenanthrene for SH044 and 45 in 04

# 2. compare distributions with SH018 samples (all months) omitted

WB_clean %>%
  filter(ID != "SH018") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() #visually confirms almost all of the outlying points belonged to SH018

# 3. compare distributions with SH018 samples (all months) and SH048_06 omitted

WB_clean %>%
  filter(ID != "SH018",
         ID != "SH048") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10(breaks = c(1,10,100,1000),    
    limits = c(0.1,900)) +
  theme_bw() 

# 4. Dig into the questionnaire data for these participants a little

part %>%
  filter(ID == "SH018") # No obvious explanation. Does not smoke, old house (1930's), reported odor, has AC, no air freshener, gas stove, retired. Look at location.

```

# Coefficient of Variation for 30 or more detections

```{r}
(thirty_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>29)) 

chemical_list30 <- as.vector(thirty_dets$ParameterName) # save the list of analytes with 30 or more detections 
WB_clean_thirty_dets <- subset(WB_clean, ParameterName %in% chemical_list30)


WB_clean_thirty_dets %>% #by month; group 1 much more variable
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #group 1
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_thirty_dets %>% #group 2
  filter(Month == 3|
           Month == 5| 
           Month == 7|
           Month == 9) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #by ID
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = ID)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #boxplot by ID - individual CVs
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #all samples combined
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()
```

# Coefficient of Variation for 30 or more detections - no SH018

```{r}
(thirty_dets <- WB_clean %>%
   filter(ID != "SH018") %>%
  count(ParameterName) %>%
  filter(n>29)) 

chemical_list30 <- as.vector(thirty_dets$ParameterName) # save the list of analytes with 30 or more detections 
WB_clean_thirty_dets <- subset(WB_clean, ParameterName %in% chemical_list30)

WB_clean_thirty_dets %>% #by month; group 1 much more variable
  filter(ID != "SH018") %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #group 1
  filter(ID != "SH018") %>%
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n=n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_thirty_dets %>% #boxplot by ID - individual CVs
  filter(ID != "SH018") %>%
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #combined
  filter(ID != "SH018") %>%
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()


#integrate chemical parameters in the dataset to sort chemicals in the y-axis
(CV_Name <- WB_clean_thirty_dets %>% #create names to match the chem_info data
  filter(ID != "SH018") %>%
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  rename("Name" = ParameterName))

(cheminfo_30WBdets <- subset(chem_info, Name %in% chemical_list30))
(CV_cheminfo <- left_join(CV_Name, cheminfo_30WBdets, by ="Name")) 

#relationship between MW and CV
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, MW), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

#show rings and CV (organized by M.W.)
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, MW), color = n)) +
  geom_point(aes(size = Rings)) + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

#relationship between koa (octanol, air partitioning coefficient) and CV
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, `log Koa`), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()
```

# Coefficient of Variation for 10 or more detections

```{r}
(ten_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>9)) 

chemical_list10 <- as.vector(ten_dets$ParameterName) # save the list of analytes with 10 or more detections (37 analytes)
WB_clean_ten_dets <- subset(WB_clean, ParameterName %in% chemical_list10)


WB_clean_ten_dets %>% #by month; group 1 much more variable
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #group 1
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_ten_dets %>% #group 2
  filter(Month == 3|
           Month == 5| 
           Month == 7|
           Month == 9) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #by ID
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = ID)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #boxplot by ID - individual CVs
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #combined
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

```

# First look at time outside and open windows

```{r}
freshair <- monthly_meta %>%
  separate(Sample, c("ID", "Type", "Month")) %>%
  filter(Type == "WB") %>%
  mutate(Month = as.numeric(Month))

#calculate sum PAHs
WB_sum_PAH <- WB_clean %>%
  group_by(ID, Month) %>%
  summarize(sumPAH = sum(Result), nPAH = n())

freshair_WB <- left_join(WB_sum_PAH, freshair, by =c("Month", "ID"))
freshair_WB

#first, how did proportion of time spent outside change? Omitting NAs
freshair_WB %>%
  filter(Outside_bin != "NA") %>%
  ggplot(aes(x = Month, fill = Outside_bin)) +
  geom_bar(position = "fill") 

#With windows open? Omitting NAs
freshair_WB %>%
  filter(Window_bin != "NA") %>%
  ggplot(aes(x = Month, fill = Window_bin)) +
  geom_bar(position = "fill") 
```

# Windrose

```{r}
library("clifro")

weather$Month <- as.factor(weather$Month)
windrose(speed = weather$AvgWindSp_mph, direction = weather$AvgWindDir, facet = weather$Month, n_col = 3)
windrose(speed = weather$AvgWindSp_mph, direction = weather$AvgWindDir)
```

# Summary Table - Questionnaire Responses

```{r}
#filter out participants that didn't return any samples
  
part_trim <- part %>%
  filter(ID != "SH022",
         ID != "SH023") 

t <- tplyr_table(part_trim, Group) %>% #could add cols = Type to further separate into air and WB-only participants; or remove "Group" for combined
  add_total_group()

# general demographics
l1 <- group_count(t, Age, by = "Age")
l2 <- group_count(t, Gender, by= "Gender")
l3 <- group_count(t, Race, by= "Race")
l4 <- group_count(t, Ethnicity, by= "Ethnicity")
l5 <- group_count(t, Education, by= "Education")
l6 <- group_count(t, Income, by= "Income")

dem <- t %>%
  add_layers(l1, l2, l3, l4, l5, l6) %>% #can add layers here as a list
  build() %>%
  kable()
 
write.table(dem, "demographics.csv")

#housing characteristics
h1 <- group_count(t, HouseAgeBin, by= "HouseAge")
h2 <- group_count(t, Odor, by= "Odor")
h3 <- group_count(t, AC, by= "AC")
h4 <- group_count(t, Flooring, by= "Flooring")
h5 <- group_count(t, Garage, by= "Garage")

#home cleaning, combustion, and fragrances
c1 <- group_count(t, Vacuum, by= "Vacuum")
c2 <- group_count(t, Mop, by= "Mop")
c3 <- group_count(t, vars(AirFreshener, AFCount), by= "AirFresh")
c4 <- group_count(t, Candles, by= "Candles")
c5 <- group_count(t, vars(AirFilter, AirFilterFreq), by= "AirFilter")
c6 <- group_count(t, vars(Fireplace, FireplaceTemp), by= "Fireplace")
c7 <- group_count(t, BackyardBurn, by= "BackyardBurn")
c8 <- group_count(t, WoodStove, by= "WoodStove")
c9 <- group_count(t, GasStove, by= "GasStove")
c10 <- group_count(t, Grill, by= "Grill")

#other habits and behaviors
b1 <- group_count(t, Bonfires, by= "Bonfires")
b2 <- group_count(t, Time_away, by= "Time_away")
b3 <- group_count(t, vars(Smoke, SmokeLoc), by= "Smoking") #smoke is the outside var and smokeloc is a subset (includes total for smokers + loc)
b4 <- group_count(t, vars(Smoke, SmokeType), by= "SmokeType") #redundant, but I want the subset I think
b5 <- group_count(t, vars(SmokeSH, SmokeLocSH), by= "SmokingSH")
b6 <- group_count(t, vars(Machinery, MachineryType), by= "Machinery")
b7 <- group_count(t, PAHOccupation, by = "PAHOccupation")

home_behaviors <- t %>%
  add_layers(h1,h2,h3,h4,h5,c1,c2,c3,c4,c5,c6,c7,c8,c9,c10,b1,b2,b3,b4,b5,b6,b7) %>% #can add layers here as a list
  build() %>%
  kable()
 
write.table(home_behaviors, "home_behaviors.csv")

#table for monthly outdoor and window data

(Freshair_grp <- freshair %>%
  mutate(Group = ifelse(Month %in% c(2,4,6,8), "1", "2")))

f <- tplyr_table(Freshair_grp, Month) %>% 
  add_total_group()

#layers
o1 <- group_count(f, Outside_bin, by= "Outside")
o2 <- group_count(f, Window_bin, by= "Window_open")

fresh_table <- f %>%
  add_layers(o1,o2) %>% 
  build() %>%
  kable()

write.table(fresh_table, "Fresh_Air.csv")

```

# Summary Table - Wristband PAH Data

```{r}
#WB data currently in ng/g
WBD
(WB_clean <- WBD %>%
  filter(Name != "SH044_WB_08",
         Name != "SH045_WB_08") %>% # samples to omit
   filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  separate(Name, into = c("ID", "Type", "Month")) %>%
  select(ID, Month, ParameterName, Result, Flag) %>%
  mutate(Flag = replace_na(Flag, "D")) %>% #D represents detections
  filter(Flag != "U"))

(WB_clean_44 <- WBD %>% #don't omit 44 and 45
   filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  separate(Name, into = c("ID", "Type", "Month")) %>%
  select(ID, Month, ParameterName, Result, Flag) %>%
  mutate(Flag = replace_na(Flag, "D")) %>% #D represents detections
  filter(Flag != "U"))

# This data excludes all U flags (non-detects) but includes all J-flags; SH044_WB_08 and SH045_WB_08 are excluded; analytes benz[a]anthracene, benz[j] and [e]aceanthrylene, 2,6-dimethylnaphthalene, and 2,6-diethylnaphthalene are excluded.

# Samples per time point (160 total):
WB_clean %>%
  select(ID, Month) %>%
  distinct() %>%
  group_by(Month) %>%
  summarize(n())

# Summary table (by month):
WB_table <- WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), geom_mean = exp(mean(log(Result))), geoSD = geoSD(Result), median = median(Result), sd = sd(Result), n = n(), min = min(Result), max = max(Result), Q1 = quantile(Result, 0.25), Q3 = quantile(Result, 0.75)) %>% 
  mutate(se = sd/sqrt(n)) %>%
  mutate(det_freq = n/max(n)) %>% #at least one analyte is detected in all samples in all months, so max(n) = the number of samples in the group
  dplyr::select(-sd) %>%
  pivot_wider(names_from = Month, values_from = c(n, det_freq, mean, se, geom_mean, geoSD, min, Q1, median, Q3, max), names_vary = "slowest")

WB_table <- WB_table[order(rev(WB_table$n_02)),]
WB_table

#50 chemicals detected across all timepoints

write_delim(WB_table, file = "WB_Data.csv", delim = "@")

# Summary of flags in each of these months as well
WBflag_table <- WB_clean %>%
  select(ParameterName, Month, Flag) %>%
  group_by(Month, ParameterName, Flag) %>%
  summarize(Count = n()) %>%
  pivot_wider(names_from = Flag, values_from = Count, values_fill = 0) %>%
  mutate(J = J + JB) %>% # JB flags are both flags applied to the same sample, this is adding these to total J-flags.
  select(-JB) %>%
  pivot_wider(names_from = Month, values_from = c(B,J), names_vary = "slowest", values_fill = 0)

write_delim(WBflag_table, file = "WBflag_Data.csv", delim = "@")


# Summary table across all months
# Numeric
WB_table2 <- WB_clean %>%
  group_by(ParameterName) %>%
  summarize(mean = mean(Result), geom_mean = exp(mean(log(Result))), geoSD = geoSD(Result), median = median(Result), sd = sd(Result), n = n(), min = min(Result), max = max(Result), Q1 = quantile(Result, 0.25), Q3 = quantile(Result, 0.75)) %>% 
  mutate(se = sd/sqrt(n)) %>%
  mutate(det_freq = n/max(n)) %>% #at least one analyte is detected in all samples in all months, so max(n) = the number of samples in the group
  dplyr::select(-sd) 

WB_table2 <- WB_table2[order(rev(WB_table2$n)),]
write_delim(WB_table2, file = "WB_Data2.csv", delim = "@")

# Summary of flags 
WBflag_table2 <- WB_clean %>%
  select(ParameterName, Flag) %>%
  group_by(ParameterName, Flag) %>%
  summarize(Count = n()) %>%
  pivot_wider(names_from = Flag, values_from = Count, values_fill = 0) %>%
  mutate(J = J + JB) %>% # JB flags are both flags applied to the same sample, this is adding these to total J-flags.
  select(-JB) 

WBflag_table2 <- WBflag_table2[order(rev(WBflag_table2$D)),]
write_delim(WBflag_table2, file = "WBflag_Data2.csv", delim = "@")
