---
title: "St. Helens Data"
output: html_document
date: "2024-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(ggplot2)
library(rio)
library(ggExtra)
library(ggcorrplot)
library(psych)
```

# log of changes 

# 1/27 to 2/12/2024:
## changed order of the heatmaps to be consistent (all by det freq) 
## corrected the list of analytes used in the correlation plot to ensure there are 5 or more pairs of observations per square
## created a variety of plots using coefficients of variation with at least 10 and 30 detections

# 2/12 - 2/26/2024
## added coefficient of variation colors to indicate number of detections considered

# Analyte omissions

## benz[a]anthracene
## benz[j] and [e]aceanthrylene
## 2,6-dimethylnaphthalene
## 2,6-diethylnaphthalene


# Problematic samples

## Reccommend Sensitivity Analysis
### SH019_A_03
### SH007_A_06

## Omit
### SH044_WB_08
### SH045_WB_08


# Load in all data from the deidentified data book 

```{r}
Data_list <- import_list("SHAirData_Deidentified.xlsx")
```

# Rename each sheet as individual dataframes

```{r}
monthly_meta <- Data_list$monthly_meta
part <- Data_list$participants
comm_coord <- Data_list$community_samples
WBD <- Data_list$chemical_data_WB
AirD <- Data_list$chemical_data_Air
EnvD <- Data_list$environmental_data
```

# Clean Up WB Data

```{r}
#WB data currently in ng/g
(WB_clean <- WBD %>%
  filter(Name != "SH044_WB_08",
         Name != "SH045_WB_08") %>% #samples to omit
   filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  separate(Name, into = c("ID", "Type", "Month")) %>%
  select(ID, Month, ParameterName, Result, Flag) %>%
  mutate(Flag = replace_na(Flag, "D")) %>% #D represents detections; easier to work with than NA values
  filter(Flag != "U"))
```


# Data Visualization - data distributions and heatmaps

```{r}
# concentration distributions by chemical and month
WB_clean %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() 

# median ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(median = median(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = median)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed() 

# mean ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = mean)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()

# detection frequency
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  ggplot(aes(y=reorder(ParameterName, det_freq), x=Month, fill = det_freq)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c() +
  coord_fixed()

```

# Stacked bar chart - median concentrations (>50% rate of detection within each month)

```{r}
#stacked bar chart; limit to 50% detection rate
#23 chemicals detected in at least 50%

pal <- c("#5A5156", "#E4E1E3", "#1C7F93", "#1CBE4F", "#B5EFB5", 
           "#3283FE", "#FEAF16", "#B00068", "#1CFFCE", "#90AD1C", 
           "#DEA0FD", "#F8A19F", "#325A9B", "#C4451C", "#1C8356", 
           "#85660D", "#FBE426", "#683B79", "#822E1C", 
           "#F7E1A0", "#C075A6", "#BDCDFF", 
           "#7ED7D1", "#2ED9FF", "#66B0FF", "#3B00FB")

# stacked means (ng/g)
WB_clean %>% #this is mean concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>% #at least 50% detection rate
  ggplot(aes(x=Month, y= mean, fill = reorder(ParameterName, mean))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 

# stacked medians (ng/g)
WB_clean %>% #this is median concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), median = median(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>%
  ggplot(aes(x=Month, y= median, fill = reorder(ParameterName, median))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 
```

# Correlation of chemicals

```{r}
# First convert data to a matrix with "NA" values for BLOD instances where each row is an analyte (5+ detections) and each row is a different WB sample (at least one detection).

(five_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>4)) 

chemical_list <- as.vector(five_dets$ParameterName) # save the list of analytes with 5 or more detections (45 analytes)
WB_clean_five_dets <- subset(WB_clean, ParameterName %in% chemical_list) # create a new WB dataset with only the list of analytes saved above

WB_clean_wide <- WB_clean_five_dets %>%
  select(-Flag) %>%
  pivot_wider(names_from = ParameterName, values_from = Result)

WB_clean_wide <- as.data.frame(WB_clean_wide)

row.names(WB_clean_wide) <- paste(WB_clean_wide$ID, WB_clean_wide$Month, sep="_")

WB_mat_ready <- WB_clean_wide %>%
  select(-ID, -Month)

(WB_mat <- as.matrix(WB_mat_ready))

# Identify "problem" chemicals that have less than 5 pairs of data for a comparison

problems <- pairwiseReport(WB_mat, diagonal = F, cut = 5)
print(problems, short=F)

problem_list_rows <- c("2-methylanthracene", "6-methylchrysene", "anthracene", "benzo[a]fluorene", "benzo[a]pyrene", "benzo[b]fluorene", "benzo[ghi]perylene", "benzo[j]fluoranthene", "benzo[k]fluoranthene", "coronene", "cyclopenta[cd]pyrene", "indeno[1,2,3-cd]pyrene", "perylene")

#removing chemicals with fewer than 5 pairs of observations from the matrix
(five_pairs <- WB_clean_five_dets %>%
    filter(!ParameterName %in% problem_list_rows)) 

# log10 transform data and shape into a matrix for correlation plot
(logWB_clean_wide <- five_pairs %>%
  mutate(log10Result = log10(Result)) %>% 
  select(-Flag, -Result) %>%
  pivot_wider(names_from = ParameterName, values_from = log10Result))

logWB_clean_wide <- as.data.frame(logWB_clean_wide)
row.names(logWB_clean_wide) <- paste(logWB_clean_wide$ID, logWB_clean_wide$Month, sep="_")
(logWB_mat_ready <- logWB_clean_wide %>%
  select(-ID, -Month))

(logWB_mat <- as.matrix(logWB_mat_ready))

# Create a correlation matrix and plot

logWB_chem_corr <- cor(logWB_mat, use="pairwise.complete.obs")
logWB_pmat <- cor_pmat(logWB_mat,
  method = "pearson",
  alternative = "two.sided",
  conf.level = 0.95)

ggcorrplot(logWB_chem_corr, type = "lower") #simple
ggcorrplot(logWB_chem_corr, hc.order = TRUE, p.mat = logWB_pmat) # uses clustering to re-order chemicals; can set "lab = T" to see numbers; x's show non sig correlation

```

# Quick Look into high concentrations (looks like mostly in group 1)

```{r}
# 1. filter data by any concentrations above 500 and 1000 ng/g

WB_clean %>% 
  filter(Result > 1000) #SH018 in 02 and 04 - verify data, followed by SH048_06
WB_clean %>% 
  filter(Result > 500) #SH018 in 06 and 08; SH048_06 again; phenanthrene for SH044 and 45 in 04

# 2. compare distributions with SH018 samples (all months) omitted

WB_clean %>%
  filter(ID != "SH018") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() #visually confirms almost all of the outlying points belonged to SH018

# 3. compare distributions with SH018 samples (all months) and SH048_06 omitted

WB_clean %>%
  filter(ID != "SH018",
         ID != "SH048") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10(breaks = c(1,10,100,1000),    
    limits = c(0.1,900)) +
  theme_bw() 

# 4. Dig into the questionnaire data for these participants a little

part %>%
  filter(ID == "SH018") # No obvious explanation. Does not smoke, old house (1930's), reported odor, has AC, no air freshener, gas stove, retired. Look at location.

```

# Coefficient of Variation for 30 or more detections

```{r}
(thirty_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>29)) 

chemical_list30 <- as.vector(thirty_dets$ParameterName) # save the list of analytes with 30 or more detections 
WB_clean_thirty_dets <- subset(WB_clean, ParameterName %in% chemical_list30)


WB_clean_thirty_dets %>% #by month; group 1 much more variable
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #group 1
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_thirty_dets %>% #group 2
  filter(Month == 3|
           Month == 5| 
           Month == 7|
           Month == 9) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #by ID
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = ID)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #boxplot by ID - individual CVs
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #all samples combined
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()
```

# Coefficient of Variation for 30 or more detections - no SH018

```{r}
(thirty_dets <- WB_clean %>%
   filter(ID != "SH018") %>%
  count(ParameterName) %>%
  filter(n>29)) 

chemical_list30 <- as.vector(thirty_dets$ParameterName) # save the list of analytes with 30 or more detections 
WB_clean_thirty_dets <- subset(WB_clean, ParameterName %in% chemical_list30)

WB_clean_thirty_dets %>% #by month; group 1 much more variable
  filter(ID != "SH018") %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #group 1
  filter(ID != "SH018") %>%
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n=n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_thirty_dets %>% #boxplot by ID - individual CVs
  filter(ID != "SH018") %>%
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_thirty_dets %>% #combined
  filter(ID != "SH018") %>%
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()


#integrate chemical parameters in the dataset to sort chemicals in the y-axis
(CV_Name <- WB_clean_thirty_dets %>% #create names to match the chem_info data
  filter(ID != "SH018") %>%
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  rename("Name" = ParameterName))

(cheminfo_30WBdets <- subset(chem_info, Name %in% chemical_list30))
(CV_cheminfo <- left_join(CV_Name, cheminfo_30WBdets, by ="Name")) 

#relationship between MW and CV
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, MW), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

#show rings and CV (organized by M.W.)
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, MW), color = n)) +
  geom_point(aes(size = Rings)) + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

#relationship between koa (octanol, air partitioning coefficient) and CV
ggplot(CV_cheminfo, aes(x = CV, y = reorder(Name, `log Koa`), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()
```

# Coefficient of Variation for 10 or more detections

```{r}
(ten_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>9)) 

chemical_list10 <- as.vector(ten_dets$ParameterName) # save the list of analytes with 10 or more detections (37 analytes)
WB_clean_ten_dets <- subset(WB_clean, ParameterName %in% chemical_list10)


WB_clean_ten_dets %>% #by month; group 1 much more variable
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
  geom_vline(xintercept = 100) +
  facet_grid(~ Month) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #group 1
  filter(Month == 2|
           Month == 4| 
           Month == 6|
           Month == 8) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw() 

WB_clean_ten_dets %>% #group 2
  filter(Month == 3|
           Month == 5| 
           Month == 7|
           Month == 9) %>%
  group_by(ParameterName, Month) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() +
    geom_vline(xintercept = 100) +
  facet_grid(~Month,) + 
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #by ID
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = ID)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #boxplot by ID - individual CVs
  group_by(ParameterName, ID) %>%
  summarize(CV = (sd(Result)/mean(Result))*100) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV))) +
  geom_boxplot() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

WB_clean_ten_dets %>% #combined
  group_by(ParameterName) %>%
  summarize(CV = (sd(Result)/mean(Result))*100, n = n()) %>%
  filter(!is.na(CV)) %>%
  ggplot(aes(x = CV, y = reorder(ParameterName, CV), color = n)) +
  geom_point() + 
    geom_vline(xintercept = 100) +
  labs(x = "Coefficient of Variation (%)", y = NULL) +
  theme_bw()

```

