---
title: "St. Helens Data"
output: html_document
date: "2024-01-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(ggplot2)
library(rio)
library(ggExtra)
library(ggcorrplot)
```

# Analyte omissions

## benz[a]anthracene
## benz[j] and [e]aceanthrylene
## 2,6-dimethylnaphthalene
## 2,6-diethylnaphthalene


# Problematic samples

## Reccommend Sensitivity Analysis
### SH019_A_03
### SH007_A_06

## Omit
### SH044_WB_08
### SH045_WB_08


# Aims

## 1.	Ascertain local outdoor features that contribute to personal PAH exposure. 
##	2.  Explore the influence of demographics and personal habits on total PAH exposure.
## 3.	Compare changes in outdoor air to changes in personal PAH exposures.

# Hypotheses

## 1.	PAH quantities and profiles captured in personal silicone wristband samplers depend on their housing characteristics, their time-activity patterns, and outdoor PAH contributions.
## 2.	Changes in outdoor vapor-phase air PAHs proportionally influences PAHs found in personal silicone wristband samples.


# Load in all data from the deidentified data book 

```{r}
Data_list <- import_list("SHAirData_Deidentified.xlsx")
```

# Rename each sheet as individual dataframes

```{r}
monthly_meta <- Data_list$monthly_meta
part <- Data_list$participants
comm_coord <- Data_list$community_samples
WBD <- Data_list$chemical_data_WB
AirD <- Data_list$chemical_data_Air
EnvD <- Data_list$environmental_data
```

# Clean Up WB Data

```{r}
#WB data currently in ng/g
(WB_clean <- WBD %>%
  filter(Name != "SH044_WB_08",
         Name != "SH045_WB_08") %>% #samples to omit
   filter(ParameterName != "benz[a]anthracene",
          ParameterName != "benz[j] and [e]aceanthrylene",
          ParameterName != "2,6-dimethylnaphthalene",
          ParameterName != "2,6-diethylnaphthalene") %>% #analytes to omit
  separate(Name, into = c("ID", "Type", "Month")) %>%
  select(ID, Month, ParameterName, Result, Flag) %>%
  mutate(Flag = replace_na(Flag, "D")) %>% #D represents detections; easier to work with than NA values
  filter(Flag != "U"))
```


# Data Visualization - data distributions and heatmaps

```{r}
# concentration distributions by chemical and month
WB_clean %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() 

# median ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(median = median(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, median), x=Month, fill = median)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()

# mean ng/g heatmap
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result)) %>%
  ggplot(aes(y=reorder(ParameterName, mean), x=Month, fill = mean)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c(trans = "log10") +
  coord_fixed()

# detection frequency
WB_clean %>%
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>% #calculate detection frequency
  ggplot(aes(y=reorder(ParameterName, mean), x=Month, fill = det_freq)) + #sorted by detections
  geom_tile( color = "white", lwd = 0.2, linetype = 1) +
  scale_fill_viridis_c() +
  coord_fixed()

```

# Data Visualization - stacked bar chart

```{r}
#stacked bar chart; limit to 50% detection rate
#23 chemicals detected in at least 50%

pal <- c("#5A5156", "#E4E1E3", "#1C7F93", "#1CBE4F", "#B5EFB5", 
           "#3283FE", "#FEAF16", "#B00068", "#1CFFCE", "#90AD1C", 
           "#DEA0FD", "#F8A19F", "#325A9B", "#C4451C", "#1C8356", 
           "#85660D", "#FBE426", "#683B79", "#822E1C", 
           "#F7E1A0", "#C075A6", "#BDCDFF", 
           "#7ED7D1", "#2ED9FF", "#66B0FF", "#3B00FB")

# stacked means (ng/g)
WB_clean %>% #this is mean concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>% #at least 50% detection rate
  ggplot(aes(x=Month, y= mean, fill = reorder(ParameterName, mean))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 

# stacked medians (ng/g)
WB_clean %>% #this is median concentrations (ng/g) not sum!!
  group_by(Month, ParameterName) %>%
  summarize(mean = mean(Result), median = median(Result), sd = sd(Result), n = n()) %>%
  mutate(det_freq = n/max(n)) %>%
  filter(det_freq > .5) %>%
  ggplot(aes(x=Month, y= median, fill = reorder(ParameterName, median))) +
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_discrete(type  = pal) +
  theme_bw() 

```

# Data Visualization - Correlation of chemicals

```{r}
# First convert data to a matrix with "NA" values for BLOD instances where each row is an analyte (5+ detections) and each row is a different WB sample (at least one detection).

(five_dets <- WB_clean %>%
  count(ParameterName) %>%
  filter(n>4)) 

chemical_list <- as.vector(five_dets$ParameterName) # save the list of analytes with 5 or more detections (45 analytes)

WB_clean_five_dets <- subset(WB_clean, ParameterName %in% chemical_list) # create a new WB dataset with only the list of analytes saved above

WB_clean_wide <- WB_clean_five_dets %>%
  select(-Flag) %>%
  pivot_wider(names_from = ParameterName, values_from = Result)

WB_clean_wide <- as.data.frame(WB_clean_wide)
row.names(WB_clean_wide) <- paste(WB_clean_wide$ID, WB_clean_wide$Month, sep="_")
WB_mat_ready <- WB_clean_wide %>%
  select(-ID, -Month)

(WB_mat <- as.matrix(WB_mat_ready))

# Create a correlation matrix and plot

WB_chem_corr <- cor(WB_mat, use="pairwise.complete.obs")
ggcorrplot(WB_chem_corr, type = "lower") #simple
ggcorrplot(WB_chem_corr, hc.order = TRUE, type = "lower") # uses clustering to re-order chemicals

```


# Quick Look into outliers (looks like mostly in group 1)

```{r}
# 1. filter data by any concentrations above 500 and 1000 ng/g

WB_clean %>% 
  filter(Result > 1000) #SH018 in 02 and 04 - verify data, followed by SH048_06
WB_clean %>% 
  filter(Result > 500) #SH018 in 06 and 08; SH048_06 again; phenanthrene for SH044 and 45 in 04

# 2. compare distributions with SH018 samples (all months) omitted

WB_clean %>%
  filter(ID != "SH018") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10() +
  theme_bw() #visually confirms almost all of the outlying points belonged to SH018

# 3. compare distributions with SH018 samples (all months) and SH048_06 omitted

WB_clean %>%
  filter(ID != "SH018",
         ID != "SH048") %>%
  ggplot(aes(y=reorder(ParameterName, Result), x=Result)) +
  geom_point() +
  facet_grid(cols = vars(Month)) +
  scale_x_log10(breaks = c(1,10,100,1000),    
    limits = c(0.1,900)) +
  theme_bw() 

# 4. Dig into the questionnaire data for these participants a little

part %>%
  filter(ID == "SH018") # No obvious explanation. Does not smoke, old house (1930's), reported odor, has AC, no air freshener, gas stove, retired. Look at location.

```
